### 2016-02-08 gd358fit.tex
Made changes to msnew.tex

\begin{deluxetable}{lllll}
\tablecolumns{5}
\tablewidth{0pc}
\tablecaption{List of 15 Periods Used in GD358 Fit and Best Fit Periods.  
\label{gd}
}
\tablehead{

\colhead{$k$} & \colhead{Frequency}& \colhead{Period} & \colhead{Uncertainty} & \colhead{Best fit periods}  \\
\colhead{(\muHz)}  & \colhead{(s)} & \colhead{(\muHz)} & \colhead{(s)} & \colhead{(s)} \\
}
\startdata

8  &2363.318 & 423.13 & 0.04 & 421.08 \\
9  &2155.544 & 463.92 & 0.04 & 458.11 \\
10 &2007.628 & 497.83 & 0.2 & 498.16 \\
11 &1857.716 & 538.30 & 0.3 & 535.44 \\
12 &1741.505 & 574.22 & 0.1 & 583.28 \\
13 &1619.700 & 617.40 & 0.2 & 616.91 \\
14 &1518.160 & 658.69 & 0.5 & 658.87 \\
15 &1428.943 & 699.82 & 0.2 & 701.53 \\
16 &1369.336 & 730.28 & 0.8 & 735.98 \\
17 &1299.147 & 769.74 & 0.2 & 771.93 \\
18 &1238.129 & 807.67 & 0.2 & 806.54 \\
19 &1170.207 & 854.55 & 0.6 & 858.64 \\
20 &1109.271 & 901.49 & 0.8 & 898.45 \\
22 &1032.967 & 968.09 & 1.4 & 970.54 \\
24 &941.333  & 1062.32 & 3.1 & 1054.75\\
\hline
 &  &  & & $\sigma_{\rm{rms}}$   & 4.22~s\\
\tableline
\enddata
\tablecaption{A list of the 15 periods used in the present fit of GD 358. We also include the corresponding periods 
from the best fit model (section ref{}). The $k$ identification is listed for each mode. There is strong evidence that 
all modes belong to an $\ell=1$ sequence. We assume that all the periods listed correspond to the $m=0$ 
component of any multiplet.\label{t1}}
\end{deluxetable}