### 2016-02-13 msnew.tex
Updated some figures. Checked all references and other formatting issues. Some revisions to text all over. New discussion and conclusion section.

%\documentclass[manuscript,natbib]{aastex}
\documentclass[12pt,preprint]{aastex}

%%\newcommand{\vdag}{(v)^\dagger}
\newcommand{\bvf}{Brunt-V\"ais\"al\"a }
\newcommand{\teff}{$\rm T_{eff}$}
\newcommand{\logg}{$\log$g }
\newcommand{\kic}{KIC }
\newcommand{\kicnsp}{KIC}
\newcommand{\sigrms}{$\sigma_{\rm RMS}$}
\newcommand{\mstar}{$M_*$}
\newcommand{\menv}{$M_{\rm env}$}
\newcommand{\mhe}{$M_{\rm He}$}
\newcommand{\xo}{$X_o$}
\newcommand{\qfm}{$q_{\rm fm}$}
\newcommand{\muHz}{\mbox{$\mu$Hz}}
\newcommand{\msun}{$M_\odot$}

\shorttitle{Asteroseismology of GD358}
\shortauthors{Bischoff-Kim}
\usepackage{graphicx}
%\usepackage{array} %Don't activate - doesn't play nice with deluxe tables!!!
\usepackage{epsfig}
\usepackage[outdir=./]{epstopdf}
%\usepackage{ulem}
%\usepackage{booktab}
%\usepackage{natbib}

%\bibliographystyle{natbib}

\begin{document}

\title{Asteroseismology of GD358 with complex C/O core profiles}

\author {Agn\`es~Bischoff-Kim\altaffilmark{1},
J.~L. Provencal\altaffilmark{2,3},
M.~H.~Montgomery\altaffilmark{3,4},
Samuel~T.~Harrold\altaffilmark{4},
B.~Howard\altaffilmark{3},
W.~Strickland\altaffilmark{5},
D.~Chandler\altaffilmark{5},
D.~Campbell\altaffilmark{5},
A.~Arredondo\altaffilmark{5},
R.~Linn\altaffilmark{5},
D.~P.~Russell\altaffilmark{5},
D.~Doyle\altaffilmark{5},
A.~Brickhouse\altaffilmark{5},
R.~Linn\altaffilmark{5},
D.~Peters\altaffilmark{5},
A.~V.~Kusakin\altaffilmark{6},
A.~V.~Sergeev\altaffilmark{7},
M.~Andreev\altaffilmark{7},
A.~Maksim\altaffilmark{7},
S. Velichko\altaffilmark{7}
R.~Janulis\altaffilmark{8},
F.~Ali\c cavu\c s\altaffilmark{9},
N. Horoz\altaffilmark{9},
S.~Zola\altaffilmark{10,11},
W.~Og{\l}oza\altaffilmark{10,11},
D.~Koziel-Wierzbowska{10,11},
T.~Kundera\altaffilmark{10,11},
D. Jableka\altaffilmark{10,11},
B. Debski\altaffilmark{10,11},
S. Meingast\altaffilmark{12}
}
\altaffiltext{1}{Penn State Worthington Scranton, Dunmore, PA 18512; axk55@psu.edu}
\altaffiltext{2}{University of Delaware, Department of Physics and Astronomy
Newark, DE 19716; jlp@udel.edu}
\altaffiltext{3}{Delaware Asteroseismic Research Center, Mt. Cuba Observatory,
Greenville, DE 19807}
\altaffiltext{4}{Department of Astronomy, University of Texas, Austin, TX 78712;
mikemon@rocky.as.utexas.edu}
\altaffiltext{5}{Meyer Observatory and Central Texas Astronomical Society, 209
Paintbrush, Waco, TX 76705; chandler@vvm.com}
\altaffiltext{6}{Fesenkov Astrophysical Institute, Almaty 050020, Kazakhstan}
\altaffiltext{7}{Ukrainian National Academy of Sciences, Main Astronomical
Observatory, Golosiiv, Kiev 022 252650; sergeev@terskol.com }
\altaffiltext{8}{Institute of Theoretical Physics and Astronomy, Vilnius
University, Vilnius, Lithuania; jr@itpa.lt}
\altaffiltext{9}{Ulupinar Observatory, \c Canakkale Onsekiz Mart University, Turkey}
\altaffiltext{10}{Astronomical Observatory of the Jagiellonian University, ul. Orla
171, 30-244 Cracow, Poland; szola@oa.uj.edu.pl}
\altaffiltext{11}{Mount Suhora Observatory, Cracow Pedagogical University, Ul.
Podchorazych 2, 30-084 Krakow, Poland; zola@astro1.as.ap.krakow.pl}
\altaffiltext{12}{Institut f\H ur Astrophysik, Universit\H at Wien, T\H urkenschanzstrasse 17, 1180 Wien, Austria}

\clearpage
\newpage

\begin{abstract}
We report on ......
\end{abstract}

\keywords{Stars: oscillations --- Stars: variables: general --- white dwarfs}

\section{Astrophysical Context}
\label{intro}
White dwarfs are the end product of evolution for around 98\% of the stars in our Galaxy. 
Buried in their interiors are the records of physical processes that take place during 
earlier stages in the life of the star. Nuclear reaction rates during the core helium 
burning phase set the core composition of white dwarfs, while the relative time spent 
burning hydrogen and helium during the asymptotic-giant-branch (AGB) phase and mass-loss 
episodes determine the thickness of the helium layer \citep{Lawlor06,Althaus05}. Helium 
atmosphere white dwarfs (DBs) comprise roughly 20\% of the population of field white 
dwarfs, with most of the remaining 80\% consisting of their hydrogen atmosphere (DA) 
cousins. The bifurcation into two spectral classes is thought to occur during post-AGB 
evolution when, in some cases, a very late thermal pulse burns off the residual hydrogen 
in the envelope, producing a nearly pure helium atmosphere \citep{Iben83}. Such objects 
are then supposed to return to the white dwarf cooling track as PG 1159 stars, which are 
widely believed to be the precursors of most DB white dwarfs. DBs are found to pulsate at 
effective temperatures ranging between 21,000 K and 28,000 K \citep{Beauchamp99, Castanheira05}.

The subject of this paper, GD358 (V777 Her) is the brightest ($m_{\rm v}=13.7$) and best 
studied helium atmosphere white dwarf pulsator. It is located near the red edge of the 
instability strip, with \teff of 24000~K and \logg of 7.8 \citep{Koester2013}. GD358's 
pulsation spectrum contains a series of independent modes, many with complex 
frequency structure.  Models involving magnetic fields and oblique rotation have been 
proposed to explain that structure \citep{Montgomery10}.

Pulsating white dwarfs do not pulsate in all their available modes at all times.  
We have maintained an active observing program of this complex star in an effort 
to identify additional frequencies. These new observations have identified additional 
periods in GD358's pulsation spectrum, bringing the total known modes to 15. Thirteen of 
these modes belong to a consecutive $\ell=1$ sequence, the longest sequence observed 
in a DBV. This sequence is critical to constrain the asteroseismic fits and better define the
limitations of high precision white dwarf asteroseismology.

Paradoxally, among the DBVs that have enough detected periods to be fitted 
asteroseismically, GD358 is the only one that has not been analyzed using the complex 
C/O profiles adapted and parameterized from stellar evolution calculations
(e.g. {}\citet{Salaris97,Althaus05}). The most recent fits of GD358 \citep{Metcalfe03c} were 
performed using 11 observed modes and simple models where the oxygen abundance 
drops linearly from its central value to zero. We present a new detailed asteroseismic analysis of GD358, taking advantage of 
its long sequence of $\ell=1$ modes to better constrain the asteroseismic fits and 
define the limits of high precision white dwarf asteroseismology.

The present analysis allows us to place GD358 in the context of stellar evolution. 
According to the models, the precursors of DBs emerge from the born-again phase with 
envelopes containing a nearly uniform mixture of helium (He), carbon (C), and oxygen (O) 
out to the photosphere \citep{Dreizler98,Herwig99}. As the PG 1159 stars cool, the helium 
diffuses upward and gradually accumulates to form a chemically pure surface layer. Through 
the DBV instability strip, this process is still ongoing so that instead of a pure helium 
layer surrounding the carbon and oxygen core, one has a region where the carbon and helium 
are still mixed. This leads to a double layered structure, with the pure He surface layer 
overlying the remainder of the uniform He/C/O envelope, all above the degenerate C/O core. 
A key prediction of the diffusion models is that, for a given stellar mass, the pure He 
surface layer will steadily grow thicker as the DB star cools.  GD358 is the fourth DBV 
we can use to check this theory. The other three DBVs 
(\citet{Bischoff-Kim14,Sullivan08, Metcalfe03c}) have painted a picture qualitatively consistent with 
the diffusion calculations, with the hotter best fit models having thinner pure helium 
layers. With GD358, we seek to confirm this trend.

In Section \ref{data}, we present our new observations and outline the data reduction process.  
In Section \ref{freq}, we establish the framework for frequency identification, and present the 
list of frequencies used for the asteroseismic investigation. We perform further analysis of the observed frequencies in Section \ref{analysis}, and motivate the $\ell$ and $m$ identification of the modes. We present the asteroseismic fitting of GD358's pulsation spectrum in section \ref{fitting}, present the results in section \ref{results} and discuss our results in section \ref{discussion}.

\section {New Observations and Data Reduction}\label{data}

GD358 was discovered in 1982 \citep{Winget82} and has been the target of the Whole 
Earth Telescope (WET) in 1990, 1994, 2000, and 2006 \citep{Provencal09, Kepler03,Winget94}. 
New observations presented here include 254 individual observing runs (1105.1 hrs) spanning 
2007-2015 (Table \ref{journal}). Each season of observations was obtained as part of 
multisite WET campaigns.

Standard WET procedure calls for observers to transfer raw images and 
calibration files to WET headquarters for analysis at the end of each local night. CCD 
data reduction follows the steps described in \citet{Provencal12}. In brief, the new observations 
were obtained with different CCD photometers, each with distinct effective bandpasses. 
We minimize the bandpass issues by using CCDs with similar detectors whenever possible and 
employing a red cutoff filter (BG40 or S8612) to normalize wavelength response and 
reduce extinction effects. We corrected each image for bias and thermal noise, and 
normalized by the flat field. Aperture photometry using the Maestro photometry pipeline 
described by \citet{Dalessio10} was performed on each image, utilizing a range of 
aperture sizes for the target and comparison stars. For each individual nightly run, we 
chose the combination of aperture size and comparison star(s) producing the highest 
signal/noise light curve.

We used the WQED pipeline \citep{wqed} to examine each nightly light curve for photometric 
quality, remove outlying points, divide by suitable comparison stars, and correct for 
differential extinction. Our observational technique is therefore not sensitive to 
oscillations longer than a few hours. The result is a series of light curves 
with amplitude variations represented as fractional intensity (mmi). The unit is a linear 
representation of the fractional intensity of modulation (1 mmi $\approx$ 1 mmag). We present
our Fourier transforms (FTs) in units of modulation amplitude 
($1 {\rm mma}=1 \times 10^{-3} {\rm ma} = 0.1 \% = 1 {\rm ppt}$).

The final reduction step is to combine the individual light curves (one example shown in Fig. \ref{tersk}) to produce 
complete light curves for GD358 for each observing season. We assume GD358 oscillates about 
a mean light level. This assumption allows us to carefully assess overlapping segments from 
different telescopes and identify and correct any vertical offsets. As discussed in detail 
in \citet{Provencal09}, we find no significant differences between the noise levels of 
amplitude spectra using: 1) the combination of all light curves including overlapping 
segments from different telescopes, 2) the combination of light curves where we retain 
only higher signal to noise observations in overlapping segments and 3) combining all light 
curves using data weighted by telescope aperture.

\begin{figure}
\epsscale{0.7}
\plotone{terskol}
\caption{Light curve of GD358 obtained with the Peak Terskol 2.0m.  Each point corresponds 
to a 10 s exposure. The nonlinear, multiperiodic nature of this star is clearly evident.
\label{tersk}
}
\end{figure}

The result is a series of new light curves for each observing season between 2007 and 2015. 
2007 contains 8.1 hrs of observation, 2008 26.1 hrs, 2009 45 hrs, 2010 201.3 hrs, 2011 
401.1 hrs, 2012 150.5 hrs, 2013 87 hrs, 2014 184.6 hrs, and 2015 55 hrs. Our coverage is 
not complete, and this incompleteness produces spectral leakage in the amplitude spectra. 
To quantify this, we sampled single sinusoids using exactly the same times as the original 
data for each season. The resulting amplitude spectrum, or ``spectral window'', is the 
pattern produced in the FT by a single frequency. The FTs for 2010, 2011, and 2014 
are given in Fig.~\ref{fts}.

\begin{figure}
\plotone{gd358multidft}
\caption{Fourier Transforms of GD358 for the 2010, 2011, and 2014 observing seasons.
The frequency range of the observed series of $\ell=1$ modes discussed in section \ref{analysis} 
is indicated by the left arrow (red).  Peaks below $\approx$ 2400 \muHz\ are combination 
frequencies discussed in section \ref{analysis} (right arrow, blue). 
(A color version of this figure is available in the online journal.)
\label{fts}
}
\end{figure}

\section{Frequency Identification}\label{freq}

Our goal is to compile a complete list of GD358's observed independent and combination frequencies 
to be used in a comprehensive asteroseismic analysis. GD358 is known for large scale changes in 
amplitudes and small but not insignificant frequency variations on a variety of timescales 
\citep{Provencal09}.  The amplitude and frequency variations evident in Fig.~\ref{fts} will 
produce artifacts in FTs, greatly complicating any frequency analysis.  
We reanalyzed all available archival data \citep{Provencal09, Kepler03, Winget94, Winget82} 
as well as our new observations to insure continuity in methodology.  We use 
{\sl Period04} \citep{Lenz05} for Fourier analysis and nonlinear least squares fitting to 
select statistically significant peaks in each FT.  As discussed in \citet{Provencal12}, 
we adopt the criterion that a peak must have an amplitude at least four times greater than the 
average noise level in the given frequency range. We define ``noise'' as the frequency-dependent 
average amplitude after prewhitening by the dominant frequencies. This is a conservative estimate, 
as it is impossible to ensure that all of the ``real'' frequencies are removed when determining
the noise level. This is certainly true for GD358, where the peaks above $\approx$ 2500 $\mu$Hz\ 
are combination frequencies \citep{Provencal09}.  Fig.~\ref{noise} displays the average noise 
as a function of frequency for the 2007-2015 observing seasons. A similar plot for the 
archival data is presented in Fig. 3 of \citet{Provencal09}. To confirm our uncertainty 
estimates, we calculated Monte Carlo simulations using the routine provided in {\sl Period04}.  
This routine generates a set of light curves using the original times, the fitted frequencies and 
amplitudes, and added Gaussian noise. A least squares fit is performed on each simulated light 
curve, with the distribution of fit parameters giving the uncertainties. Our Monte Carlo results 
are consistent with our average noise estimates.

\begin{figure}
\plotone{noise}
\caption{A comparison of the average noise as a function of frequency for the 2007-2015 observing 
seasons.  Each data set was prewhitened by its dominant frequencies.  The noise levels for each 
season are somewhat different.  This must be taken into account during frequency analysis.  
\label{noise}
}
\end{figure}

Our frequency selection procedure involves identifying the largest amplitude resolved peak in the FT,
fitting a sinusoid at that frequency to the data set, subtracting the fit from the light curve, 
recomputing the FT, examining the residuals, and repeating the process until no significant power 
remains. This technique, known as prewhitening, must be employed with an over-abundance of caution, 
especially since we are aware of amplitude and/or frequency modulation in our data set. A detailed
discussion of the prewhitening procedure and steps taken to minimize the effects of amplitude 
modulation is given in \citep{Provencal09}. Our final identifications of independent frequencies
detected in each observing season are given in Table \ref{freq1} and Table \ref{freq2}.

\section{Frequency Analysis}\label{analysis}

\subsection{Frequencies Distribution}

Perusal of Tables \ref{freq1} and \ref{freq2} shows that GD358's observed frequencies vary in two
important ways.  First, frequencies detected in a given observing season are not found 
in all observing seasons. Asteroseismology is based on the assumption that the available pulsation
frequencies are linked to stellar structure.  Since we are fairly certain that GD358's internal structure 
does not change on the timescales of the observations, we can assume that GD358 excites 
different subsets of its available pulsations at different times. While this is a common phenomenon
seen in white dwarf pulsators, the selection mechanism remains unknown. The best way to 
determine GD358's complete set of pulsation frequencies is to combine frequency identifications 
from multiple observing seasons as outlined in \citet{Kleinman98}.

\begin{figure}
\epsscale{0.8}
%\plottwo{gd358multif_2015.eps}{gd358multi.eps}
%\centering{
%\includegraphics[width=1.0]{gd358multif_2015.eps}
%}
\plotone{gd358multif_2015.eps}
\caption{A schematic representation of GD358's pulsation modes for all available data between
1982 and 2015. Systematic patterns of distribution are evident. The bands between 2400 and 1000
\muHz\ (400 and 1000 s) are of particular importance for this work. (A color version of this 
figure is available in the online journal.)
\label{schematic}
}
\end{figure}

Fig.~\ref{schematic} presents a schematic representation of GD358's pulsation frequencies for all the 
available data. The features of asteroseismic importance are the localized bands between 900 and 2400
\muHz\ (1100 and 400 s). We interpret each band to represent a single radial overtone. The bands
themselves illustrate the second way in which GD358's frequencies vary. Each band consists of 
frequency detections from multiple observing seasons, but each band is significantly wider in 
frequency space than the frequency error of any single measurement, arguing that some process 
is acting on the frequencies. The simplest explanation is rotational splitting.  However, with the 
exception of the two highest frequency (shortest period) bands ($k=8$ and $k=9$ 
from \citet{Provencal09}), we find no clear evidence of multiplet structure in the frequency 
distributions of the lower frequency (longer period) bands. The lower frequency band widths are 
most likely the result of amplitude and/or frequency modulation that obscures any underlying 
signatures of rotational splitting.

\begin{figure}
 \epsscale{1.0}
 \plotone{gd358k17_k8.eps}
% \plottwo{gd358multifk17.eps}{gd358multifk8.eps}
 \caption{The detailed schematic distribution of frequencies for the bands near 1300({\em{right}}
 and 2360 \muHz\ ({\em{left}}). The 1300 \muHz\ band shows no multiplet structure, unlike the 2360 \muHz\
 band. Here we see clear evidence of multiplet structure. Assuming the triplet structure represents
 $\ell=1$, this implies a rotation period of $\approx 1.5$ days. Note changes in the scale for both axes.
 (A color version of this figure is available in the online journal.)
 \label{rot}}
\end{figure}

Fig. \ref{schematic} also shows that the widths of the bands is not constant, but changes from band to band.
\citet{Bell15} presents an interesting analysis of the hydrogen atmosphere pulsator (DAV) KIC4552982, 
in which they identify 17 bands of pulsation frequencies. This DAV follows the general pattern exhibited 
by GD358: the highest frequency (shortest period) mode shows evidence of rotational splitting, while the 
lower frequency (longer period) modes are complex bands. The authors note that this DAV's bands have 
different widths in frequency space, and that there may be astronomical significance to this. Although the 
observational timebases and coverage are quite different (20 continuous months for KIC4552982 vs 30 
incomplete years for GD358), our data presents the opportunity to investigate this for a DB pulsator.  
We measured the width of each of GD358's band. We define ''width'' as the difference between the 
lowest and highest frequency detected in each band. Fig. \ref{thick} presents the results. 
The widths of GD358's bands are commensurate with those found in KIC4552982.  We also find a general 
increase in width with decreasing frequency (increasing period), until we reach 
the mode at 1238 \muHz\ (807 s). This band has a width at least twice as wide as any other.  What 
is different about this mode???

\begin{figure}
 \epsscale{0.8}
 \plotone{modethick.eps}
 \caption{Width of each bank of power in Fig.~\ref{schematic}. We find an overall increase 
 in band thickness with decreasing frequency (increasing period).  The band at 807~s has a width
 at least twice as large as other bands.  
 \label{thick}}
\end{figure}

\subsection{Mode Identification}
Our current work has produced a well defined sequence of modes, adding to previous studies \citep{Montgomery10, Provencal09, Metcalfe00, Winget94}. The previous identification of these modes as 
a series of $l$=1 radial overtones is based mostly on the pulsation frequency distribution 
and limited spectroscopic analysis \citep{Kotak02, Castanheira05}. It is important to further 
investigate these identifications as we initiate an indepth asteroseismic investigation.

GD358's combination frequencies provide a tool by which we can bolster $\ell$ identifications. 
Combination frequencies are typically observed in the FTs of moderate to large amplitude pulsators.  
They are identified by their exact numerical relationships with parent frequencies.  
The combinations themselves are not independent, but result from nonlinear effects associated 
with the surface convection zone \citep{Brickhill92, Brassard95, Wu01, Ising01}.  
\citet{Wu01} lays the groundwork, showing that observed amplitudes of the combination 
frequencies depend on geometric factors such as the $(\ell,m)$ indices of the parents and the 
inclination of the pulsation axis to the line of sight.

The methods outlined in \citet{Provencal12} and \citet{Montgomery10} work best when applied to 
larger amplitude frequencies detected in high signal to noise data sets such as provided by 
extensive WET runs.  The primary reason for this is that combination frequencies are lower 
amplitude than their parents, and so are more difficult to detect in sparse data.  We chose 
the 1990, 1994, 2006, 2010, 2011, and 2014 observing seasons, and looked at pulsation 
frequencies with amplitudes above 10 mma.

Fig.~\ref{modeamps} shows the resulting probability distribution of $\ell$ and $m$ values for the 1735.96 \muHz\
frequency as detected in 2014. The mode is clearly identified as an $\ell=1$, $m=1$ mode.

\begin{figure}
 \epsscale{0.8}
 \plotone{modeamps.eps}
 \caption{Probability distribution (from 0 to 1) of $\ell$ and $m$ values for the 1735.96 variation detected in 
 2014.  The solutions are from individual simulations with $Res_{rms}<9.5^{-06}$. The amplitudes of the 
 observed combination frequencies argue that this is $\ell=1$, $m$=1.  
 \label{modeamps}
 }
\end{figure}

Add table with all l,m ids
%\label{tl}
%isnt' that same as table \ref{gd}?

Finally, given the $l$=1 identification, but the lack of definite multiplet structure, the best way to 
determine the l=0 component for each band is simply to average the frequencies in the bands.  We experimented
with numerous weighting techniques, and determined there is no significant difference in our solutions.  
Our final frequency identification for each band is given in Table~\ref{gd}. We use these periods in our asteroseismic fitting.

\section{Asteroseismic fitting}
\label{fitting}

The basic method in our asteroseismic fitting consists of calculating grids of white dwarf models and 
running a fitting subroutine to match the periods of the models ($\rm P_{calc}$) with the observed 
periods and calculated residuals using the usual formula

\begin{equation}
\label{fiteq1}
\rm \sigma_{RMS} = \sqrt{\frac{\sum_{1}^{n_{obs}} {(P_{calc}-P_{obs})^2}}{n_{obs}}},
\end{equation}

\noindent where $\rm n_{obs}$ is the number of periods present in the pulsation spectrum.

\subsection{The Models}
\label{models}

To compute our models, we used the White Dwarf Evolution Code (WDEC). The WDEC evolves hot 
polytrope models from temperatures above 100{,}000~K down to the temperature of our choice. 
Models in the temperature range of interest for the present study are thermally relaxed 
solutions to the stellar structure equations. Each model we compute for our grids is the result 
of such an evolutionary sequence. The WDEC is described in detail in \citet{Lamb75} and 
\citet{Wood90}. We used smoother core composition profiles and implemented more complex profiles 
that result from stellar evolution calculations \citep{Salaris97}. We updated the envelope 
equation of state tables from those calculated by \citet{Fontaine77} to those given by 
\citet{Saumon95}. We use OPAL opacities \citep{Iglesias96} and plasmon neutrino rates 
published by \citet{Itoh96}.

DBVs are younger than their cooler cousins the DAVs. Time-dependent diffusion calculations 
\citep[e.g.][]{Dehner95,Althaus05} show that at 24{,}000~K, a typical temperature for a DBV, 
the carbon has not fully settled into the core of the star yet. We expect the helium layer to 
be separated into a mixed He/C layer with a pure He layer on top, as shown in Fig. \ref{ffit1}. 
Following \citet{Metcalfe05a}, we adopted and parameterized this structure in our models.

\subsection{Initial Grid Search}
\label{grids}

In our asteroseismic fits, we initially varied six parameters; the effective temperature, 
the mass and four structure parameters. There are two parameters associated with the shapes 
of the oxygen (and carbon) core composition profiles: the central oxygen abundance (\xo) and 
the edge of the homogeneous carbon and oxygen core (\qfm, as a fraction of stellar mass. 
For envelope structure, \menv marks the location of the base of the helium layer 
and \mhe the location where the helium abundance rises to $1$ (see Fig. \ref{ffit1}). 
\menv and \mhe are mass coordinates, defined as e.g. $M_{\rm env} = -\log(1 - M(r)/M_*)$, 
where $M(r)$ is the mass enclosed in radius r and $M_*$ is the stellar mass.

\begin{figure}
%  \includegraphics[scale=0.70,viewport=50 0 50 350]{f2.png}
%  {\includegraphics[width=1.0\columnwidth]{ffit1.eps}}
\plotone{ffit1.eps}
\caption{
{\em Upper panel:} Chemical composition profiles of the best fit model. The center of the model is to the left and the surface to the right.  $q=2$ corresponds to $M_r = 0.99 \: M_*$. The vertical axis shows fractional abundances. 
{\em Lower panel:} The corresponding \bvf frequency ($\log{N^2}$ on the vertical axis). \label{ffit1}
}
\end{figure}

We started with a master grid (table \ref{fitt1}) chosen so that it covered all relevant 
area of parameter space and had sufficient resolution to find any region of local minimum in the fitness parameter \citep{Bischoff-Kim11a,Bischoff-Kim14}. We used the minimum resolution that was computationally manageable. The master grid involved the computation 
of 10,483,200 models. We fit simultaneously all 15 periods, requiring all of them to be $\ell=1$ 
modes. A fitness map of this initial fit are shown in the left panel of Fig. \ref{ffit2}, and the parameters of the best fit model are listed in table \ref{fitt1}.

\subsection{Asymptotic Period Spacing}
\label{periodspacing}

Before we refine the period-by-period fitting optimization, it is worth stepping back and consider 
what we can learn from the average period spacing of GD358. The average period spacing provides an asteroseismic measure of the mass and temperature of the star, independent of the details of 
internal chemical composition profiles. Higher $k$ modes are not strongly trapped in the core and 
according to asymptotic theory, they should be nearly evenly spaced in period. This spacing is 
given by (Unno et al. 1989)

\begin{equation}
\label{fiteq2}
\Delta P = \frac{\pi}{\ell(\ell + 1)^{1/2}}\left[\int_{r_1}^{r_2}\frac{N}{r}dr\right]^{-1},
\end{equation}

\noindent
where $r_1$ and $r_2$ are turning points of the mode and N is the Brunt-V\"{a}is\"{a}l\"{a} frequency. 
The asymptotic period spacing is $\ell$ dependent, with higher $\ell$ modes having smaller spacing. 
In the case of GD 358, we have a single $\ell=1$ sequence so we only need to worry about the dependence 
of the asymptotic period spacing on the \bvf frequency. Much if not all of GD 358's pulsation 
spectrum is close to the asymptotic limit, because the shortest period observed is a $k=8$ mode.

The dependence of $\Delta P$ on the \bvf frequency leads to higher mass and lower temperature models 
having a smaller period spacing (their interior is less compressible). This effect appears in asteroseismic 
fitting of white dwarfs and also sdB stars as a ubiquitous diagonal trend in contour maps of the quality 
of the fits in the mass-effective temperature plane \citep[e.g][]{Bischoff-Kim14,Castanheira09,Charpinet08}. 
One requirement for the periods of the model to match the observed periods is that the average period spacing 
in the models matches the average period spacing in the observed pulsation spectrum. If a good match occurs 
for a given mass and effective temperature, then models with lower mass but higher effective temperature 
will also match well.

We can use the sequence of 13 consecutive $\ell=1$ modes found in GD358's pulsation spectrum to calibrate our models. 
Using the $\ell=1$ sequence of modes, we compute an average period spacing of 39.08 seconds. We call 
this $\Delta P_{\rm obs}$. For each model in the master grid, we compute an asymptotic period spacing 
($\Delta P_{\rm calc}$). This asymptotic period spacing is calculated by first discarding the 10 lowest $k$ modes. The exact value of 10 is somewhat arbitrary, but it is chosen so that the modes we use in 
our computation are indeed in the asymptotic limit. The higher $k$ modes show weaker trapping than the 
lower $k$ modes. We then fit a line through the set ($k_i$,$P_i$). The slope gives us the asymptotic 
period spacing in the model. We also calculate the residuals of the fits and discard the models that have residuals above a certain limit. The limit is chosen by checking the procedure 
by eye on a few models.

We show a contour map of the location of the models that best match the average period spacing of 
39.08 seconds in Fig. \ref{ffit2}, right panel. We place it side by side with a contour map showing 
the location of the best fit model in the same region of parameter space, based on the master grid 
fitting described in section \ref{grids}. Note how the best fit model falls right within the valley 
where the period spacings between GD358 and the models match. This should come as no surprise, as in 
order for 15 periods to fit reasonably well, the model period list should have a spacing similar to 
that of the observed period spectrum. It is also somewhat comforting to find that the spectroscopic 
data point also falls within the swath where the period spacings match (though the best fit model 
based on the master grid search is further down the valley).

%  \includegraphics[width=1.0\columnwidth]{ffit2.eps}
\begin{figure}
\epsscale{0.7}
\plotone{ffit2.eps}
\caption{
{\it Left panel}: Contour map showing the location of the best fit models. The quantity plotted in 
the mass-effective temperature plane is the fitness parameter defined in equation \ref{fiteq1}. 
The scale is in seconds. {\it Right panel}: Contour map of the difference between the average 
periods spacing determined from GD 358's periods and the asymptotic period spacing of the models, 
in the mass-effective temperature plane. The scale is in 10s of seconds, so the worse matches 
pictured on the plot have $\left|\Delta P_{\rm obs}- \Delta P_{\rm calc}\right| \sim 2 \; \rm s$. \label{ffit2}
}
\end{figure}

One can fit simultaneously the average period spacing and the individual periods 
formally while performing the fits by using some prescription to calculate the goodness of fit. 
This leads to a more complex relation than defined in equation \ref{fiteq1}. 
Note that the period spacing is a much weaker constraint than the individual period fit. 
If one takes 5.0 seconds as an upper limit for goodness of fit, that includes 9 points in the 
period-by-period fit plot (left panel in Fig \ref{ffit2}), but the entire parameter plane 
for the average period spacing fit plot (right panel). We bypass the difficulty of 
assigning proper weights to each constraint by taking the average period spacing fit into 
account as an initial step to refine our fits. We refine our fits by calculating models on 
a higher resolution grid that fall within the region defined by the two diagonal lines shown 
in the right panel of Fig. \ref{ffit2} (covering the effective temperature range 25,000~K-27,500~K), then performing a period-by-period fit using 
equation \ref{fiteq1}.

\subsection{Optimal Grid Resolution}
\label{refinedfits}

Having determined a more restricted region of parameter space to search for 
the best fit models, we now turn to the question of how fine we need to 
make our refined grid. We want to have a high enough resolution grid that 
we can be sure we captured a true minimum, but on the other hand, there are 
computational limitations to how many models we can afford to calculate, 
save, and process.

One way to gain a sense of how fine the grid needs to be is to make single 
parameter cuts through parameter space. Fig.  \ref{ffit3} shows such a cut 
for master grid models. The plot was made by fixing 5 of the 6 parameters 
to the best fit values of the best initial fit (see table \ref{fitt1}).
We allowed the effective temperature to vary through the entire range 
considered for the master grid. The fits seem to settle to a minimum in a 
smooth way, except for the striking anomaly at 26,600~K. Upon further 
investigation, we discovered that anomaly to be due to a single period in 
the pulsation spectrum: the 968~s mode. In the models, that mode appears 
highly sensitive to some parameters, and in a non-continuous way.

\begin{figure}
\epsscale{0.7}
\plotone{ffit3.eps}
%  \includegraphics[scale=0.70,viewport=50 0 50 350]{f2.png}
%  \includegraphics[width=0.9\columnwidth]{ffit3.eps}
\caption{
Dependence of the fitness parameter on effective temperature alone based on 
the master grid of models. \label{ffit3}
}
\end{figure}

This is unsettling, as it is easy to see that one might miss the best fit 
models if the grid is not fine enough. In order to quantify "fine enough", 
we ran systematic scans, computing models with 5 parameter fixed and allowing 
the 6th parameter to vary in very fine steps. We went down in 
step sizes to $\Delta {\rm T_{eff}}=1$~K, $\Delta M_*=0.0001 M_\odot$, 
$\Delta M_{\rm env}= \Delta M_{\rm He}=0.01$ dex, $\Delta X_o=0.01$, and $\Delta q_{\rm fm}=0.001$. 
Results for 4 different parameters are shown in Fig \ref{ffit4}. The figure 
includes zooms to regions of interest. Other modes also exhibit a behavior 
similar to the 968~s mode, in different regions of parameter space. For instance, 
the jumps in \sigrms\ for \qfm\ between 0.3 and 0.4 are due 
to sudden changes in the 854.5 and 902.5~s modes.

Regardless of the behavior of individual modes in the models as the parameters vary, 
we can assert from the test above what step sizes will allow us to minimize our risk 
of missing a minimum. We settled on optimal step sizes of 
$\Delta {\rm T_{eff}}=20$, $\Delta M_*=0.0002$, $\Delta M_{\rm He}=0.2$, 
$\Delta X_o=0.2$, and $\Delta q_{\rm fm}=0.01$ for our refined grid. To offset 
the computational cost, we fixed $M_{\rm env}$ to a canonical value of 
2.80 \citep{Dehner95}. It is clear from the second panel in Fig. \ref{ffit4} 
that there is only a weak dependence of the fits on that parameter.

\begin{figure}
\epsscale{0.9}
\plotone{ffit4.eps}
%  \includegraphics[scale=0.70,viewport=50 0 50 350]{f2.png}
%  \includegraphics[width=0.9\columnwidth]{ffit4.eps}
\caption{
\label{ffit4}
Dependence of the fitness parameter on different parameters. The 5 parameters other than the one shown on the horizontal axis in each plot are held to fixed values corresponding to the best fit model (table \ref{ffit1}).
}

\end{figure}

\section{Results of the Period Fitting}
\label{results}

The parameters for our best fit model based on the refined grid are listed in table \ref{fitt1} and the periods of that model in table \ref{gd}. The goodness of fit of the model is \sigrms$=4.22$~s. In Fig. \ref{ffit1}, we show the interior structure of the best fit model, with the corresponding \bvf frequency profile.

\subsection{Validation of the Fitting Method}

With what we have discovered in section \ref{refinedfits}, it is only natural to be concerned about whether we have truly found a best fit. We performed a simple test to validate our fitting method, which consisted of using the exact same procedure to find a best fit to the periods of a model that was not on any of the grids we calculated, but that did have parameters that were very close to the best fit model. We used the period list for a model with parameters ${\rm T_{eff}}=25570$~K, $M_*/M_\odot=0.5437$, $M_{\rm He}=-5.30$, $X_o=0.42$, and $q_{\rm fm}=0.175$, fixing $M_{\rm env}$ to -2.80, and including only the subset of 15 $\ell=1$ periods that match GD358's pulsation spectrum.

We first performed a fitting of the periods using the master (coarse) grid described in section \ref{grids}. This placed the best fit model in the appropriate region of parameter space. Then we refined our fits, using the grids described at the end of section \ref{periodspacing}. We are able to recover the best fit parameters adequately, with the top 10 best fit models having parameters ${\rm T_{eff}}=25620 \pm 67$~K, $M_*/M_\odot=0.5431 \pm 0.0011$, $M_{\rm He}=-5.32 \pm 0.098$, $X_o=0.40$ for all top 10 fits, and $q_{\rm fm}=0.175 \pm 0.5$. The top 10 best fits have \sigrms $\sim$ 0.5~s. We remind the reader that the step sizes in the second phase of the fitting are 20~K for ${\rm T_{eff}}$, 0.2 for $M_{\rm He}$, 0.2 for $X_o$, and 0.01 for $q_{\rm fm}$.

\section{Discussion and Conclusions}
\label{discussion}

We find a best fit effective temperature of 25,560~K for GD358, hotter than the recent spectroscopic determination of 24000~K \citep{Koester2013} by more than one sigma (the mass matches the spectrosocopy). For reasons discussed in section \ref{periodspacing}, while it is possible to determine whether a star is closer to the red edge or the blue edge of the instability strip, effective temperature is poorly constrained in asteroseismic fitting. As we can see from Fig. \ref{ffit3}, there are local minima that are almost as good as the global minimum we found, some at lower effective temperature. Structure parameters, especially the location of chemical transitions (with the exception of \menv\ in the case of GD358) are more tightly constrained.

In fitting GD358, we discovered that the best fit parameter space was semi-chaotic, with two very close sets of parameters able to yield very different fitness number. This is due to the fact that some periods can change values drastically for a small change in a given parameter. The phenomenon is only semi-chaotic, as the period will oscillate between two values. This is good news, because there remains a well defined minimum in parameter space, even if it is a double valued minimum. Any search algorithm that uses a small enough mesh should be able to find the minimum. This phenomenon likely became apparent while fitting GD358 because it is the first white dwarf for which we fit a long consecutive sequence of modes of the same $\ell$ with high resolution grids.

In this study, we varied the parameters that have traditionally been varied in this type of asteroseismic fitting, so that we can place our results side by side with other studies of DBVs, including KIC 8626021 \citep{Bischoff-Kim14}, EC200058 \citep{Bischoff-Kim11a}, and CBS114 \citep{Metcalfe05a}. We now have 4 DBVs that were the object of asteroseismic fitting that used a consistent set of models. Unfortunately, GD358 does not fit neatly into the trend (neither does EC20058). It is true, however that the two blue edge objects (EC20058 and KIC 8626021) have thinner helium layers ($M_{\rm He}=-6.10$ and $-7.90$ respectively than the red edge stars ($M_{\rm He}=-5.40$ for GD358 and $-5.94$ for CBS114. Pinning down the proper effective temperature of each star is challenging. To be fully consistent, it is best to use the asteroseismic values, even if effective temperature is not strongly constrained asteroseimically. At best we can contrast blue edge versus red edge objects.

Numerical experiments of the type presented in section \ref{refinedfits} \citep{Bischoff-Kim15} have shown that, consistent with the theory of non-radial oscillations, the shapes of the transition zones matter as much as where the transition zones occur. Modern asteroseismic fitting vary parameters attached to the shape of transition zones \citep{Giammichele15}. We refrained from doing this in the present study because we wanted to compare our results with previous studies of DBVs. It is unclear how much of an effect on structure parameters a change in parameterization would have.

%We have performed and published a preliminary fit of the star (Hawaii proceedings), where we found a temperature of 23,600 K, a mass of 0.605 $M_\odot$, \menv = -2.40, \mhe = -5.20, a central oxygen abundance of 0.99, and \qfm = 0.50. In the preliminary fits, the period list was not finalized and we did not include the 1065.260 s mode, because it was difficult to fit and we did not have a positive ell identification for the mode. It might be disheartening to observe such a big difference in the best fit parameters by just including or excluding one mode out of 15, but it is worth noting that the thickness of the pure helium layer was not affected much by the inclusion or exclusion of the 15th mode and refinements done in the search in this work. We emphasize again that in the final stages of our fitting for this work, we fixed the envelope mass to -2.80.
%
%In section \ref{periodspacing}, we discussed the degeneracy in the determination of the mass and effective temperature parameters. That degeneracy means that a unique determination of effective temperature and mass is not always possible.  The fits are also marginally sensitive to the central oxygen abundance. What is surprising is the big difference in the determination of the location of the edge of the homogeneous C/O core from our preliminary fits to the refined fits presented in this paper. If we take the best fit model from the preliminary work and include the 15th period when calculating the fitness number, the quality of the fit becomes significantly worse (2.84 s to 5.12 s). We would not expect a k=24 mode to be strongly trapped in the core, quite the contrary. This could be due to a core-envelope symmetry \citep{Montgomery03}, but more detailed studies would be required to make a more definitive statement. We should also keep in miind that the period lists used in the fits were close but not identical.

\acknowledgments

{\it Facilities:} \facility{MCAO:0.6m ()}, \facility{Struve()},
\facility{KPNO:2.1m ()}, 
\facility{}, \facility{BOAO:1.8m ()}, \facility{Lulin:1.8m ()},
\facility{Beijing:2.16m ()}, \facility{Maidanek:1.5m ()}, \facility{Peak Terskol}

\bibliographystyle{apj}
\bibliography{index_f}

\clearpage

\begin{deluxetable}{llrrr}
\tablecolumns{5}
\tabletypesize{\small}
\tablewidth{0pc}
\tablecaption{\label{journal}}
\tablehead{
\colhead{Run Name} & \colhead{Telescope} & \colhead{Detector} & \colhead{Date} & \colhead{Length}\\
\colhead{} & \colhead{} & \colhead{} & \colhead{} & \colhead{(hrs)}
}
\startdata
\input{journaldata.tex}
\enddata
\end{deluxetable}

\clearpage

\input{1982freq.tex} % \label{freq1}
\clearpage
\input{2011freq.tex} %\label{freq2}
\clearpage
\input{gd358fit.tex} % \label{gd}
\clearpage
\input{fitt1} % \label{fitt1}
\clearpage
%\begin{thebibliography}{}
%\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

%\end{thebibliography}

\end{document}