### 2016-03-16 gd358fit.tex
Almost final edits

\begin{deluxetable}{lllll}
\tablecolumns{5}
\tablewidth{0pc}
\tablecaption{List of 15 Periods Used in Fitting GD358 and Corresponding Best Fit Periods.  
\label{gd}
}
\tablehead{
\colhead{$k$} & \colhead{Frequency}& \colhead{Period} & \colhead{Uncertainty} & \colhead{Best fit periods}  \\
\colhead{}  & \colhead{(\muHz)} & \colhead{(s)} & \colhead{(s)} & \colhead{(s)} 
}
\startdata
8  &2363.318 & 423.13 & 0.04 & 423.12 \\
9  &2155.544 & 463.92 & 0.04 & 463.87 \\
10 &2007.628 & 497.83 & 0.2 & 493.14 \\
11 &1857.716 & 538.30 & 0.3 & 540.80 \\
12 &1741.505 & 574.22 & 0.1 & 574.90 \\
13 &1619.700 & 617.40 & 0.2 & 615.98 \\
14 &1518.160 & 658.69 & 0.5 & 656.61 \\
15 &1428.943 & 699.82 & 0.2 & 701.43 \\
16 &1369.336 & 730.28 & 0.8 & 741.67 \\
17 &1299.147 & 769.74 & 0.2 & 768.84 \\
18 &1238.129 & 807.67 & 0.2 & 809.40 \\
19 &1170.207 & 854.55 & 0.6 & 854.01 \\
20 &1109.271 & 901.49 & 0.8 & 893.00 \\
22 &1032.967 & 968.09 & 1.4 & 967.71 \\
24 &941.333  & 1062.32 & 3.1 & 1051.16\\
\hline
   &         &         & $\sigma_{\rm{rms}}$   & 0.964~s\\
   &   		 &		   & BIC (${\rm n_{obs}}=15$                 & -0.0303~s \\
\tableline
\enddata
\end{deluxetable}